Evaluating a simplified forecast model in comparison to the European forecasting hub ensemble

### 2022-09-30 paper/paper.Rmd
review introduction

Whilst there is robust and consistent evidence that ensemble forecasts provide reliable and performant forecasts across domains [@Reich2022-jo] they also have a range of downsides.
The most significant is their considerable resource cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
Forecasts from these ensembles can also be hard to interpret as it is difficult to link them to the underlying model assumptions.
Additional issues with multi-team collaborations include a lack of feedback to those contributing forecasts that would allow them to improve their forecast approaches, maintaining incentives for forecasters to continue to contribute [@Bosse2022-hm], and difficulty improving the quality of the ensemble [@Sherratt2022-ue].
Each of these may impact the long-term quality of the resulting ensemble forecasts and have implications for end-users.

The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic due to a pressing need for reliable forecasts and a perception that many publicly available forecasts were low quality. Over 2020 and 2021, teams established COVID-19 Forecasting Hubs covering the US [@Cramer2022-sp], Germany and Poland [@Bracher2021-wk], and Europe [@Sherratt2022-ue] (the latter two including authors of this study). All of these collaborations, which the author have been involved with managing, ensembled contributions from multiple independent teams using a similar approach and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting.
However, less progress has been made mitigating the downsides of these projects and improving access to the high quality and robust forecasts they seek to generate.

In this work, we draw insights from ensemble forecasts produced and endorsed by these forecast hubs, as well as our own forecasting work, to propose and evaluate a "surrogate" forecast model.
This seeks to mitigate some of the downsides of recent COVID-19 ensembles forecasts by emulating their observed behaviour, using a minimal set of easily communicated and epidemiologically justified assumptions, and limited computational resources with an easily generalised implementation.
The primary aim of this approach is to help highlight the behaviour, and potential mechanisms behind this behaviour, of ensemble forecast widely considered the gold standard for COVID-19 forecasting.
In addition to this, we seek to provide a simple, and robust, forecasting system that can be easily reused by others, perhaps where a large scale forecast hub ensemble is not available, and be used as a platform for future research both into improving understanding of ensemble behaviour and increased forecast performance.

In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysic [@Vernon2014-ov], emulation approaches have been used to circumvent some of the resource requirement issues by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components.
However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and, for ensemble models in particular, have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
This means that, whilst non-parametric emulation has utility for allowing rapid exploration of the parameter space for epidemiological models [@Iskauskas2022-fz] utility, it may be less suitable for use in replicating epidemiological ensemble forecasts than the surrogate assumption driven framework we propose.

In order to understand the behaviour of the forecast hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their own forecast models and then use them to produce weekly forecast that they then submit to the central hub team. 
Each hub targets a range of metrics including test positive reported cases, repoorted deaths, and hospitalisations, has a specific geographic focus, and asks for weekly forecasts (using epidemiological weeks i.e. Sunday to Saturday) over a time horizon of a few weeks.
Typically, the ensemble forecast produced each week comprises at least 3 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@Sherratt2022-ue].
Observed data are available and updated daily, and most submitted forecasts use this dataset, along with potentially other sources of real-time information, to produce forecasts.
Here, we focus on reported cases and primarily on the European forecast hub but our observations hold, in our view, across COVID-19 forecast hubs and to a less degree targets.
In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF; @Bosse2022-hm], and human insight based forecast models that may also include elements of other methods [@Bosse2022-hm].
Real-time evaluation has shown that each of these classes of models may perform well depending on the context and specific implementation of the forecast model [@Bosse2022-hm].

We have contributed a range of forecasts to these forecast collaborations generally focussed on semi-mechanistic statistical methods, and human insight-based forecasts.
Our forecast submissions have not clearly over- or under- performed other forecasts submitted to the forecasting hubs (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm; @Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue]).
The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
Based on our observations our forecasts have generally captured the current trend relatively well, but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns. 
Our previous methodology also requires significant computational resources running for an hour on a Azure D v5-series 16-core machine when producing forecasts for the European forecasting hub [@AzurePricing]. 
In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
We found that our human insight-based forecasts significantly outperformed our model-based forecasts on average and hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@Bosse2022-hm].

@@

Based on these insights from observing the performance of our forecast submissions and from the hub ensembles over time, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] with the aim of producing ensemble-like performance without sacrificing interpretability.
The first simplification we make is to model only weekly data, rather than using daily data and then aggregating. This mitigates the impact of daily reporting artefacts. It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain significant weight in the model.
This leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.

@@

The only observation for which we do not make an adaptation is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion.
This choice is motivated by our belief that the underlying transmission process is an exponential one and therefore a count error model, where variance is linked to the mean, is a sensible choic. 
We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations [@bracherEvaluatingEpidemicForecasts2021].

@@

```{r assemble-eval, fig.height = 12, fig.width = 12, out.width = "95%", fig.cap = "Relative weighted interval score by location, horizon, and forecast date for the surrogate forecast model compared to the ensemble forecast model on the log scale. a.) The density of the relative score by horizon. Horizontal black lines give the 5\\%, 35\\%, 65\\%, and 95\\% quantiles. b.) The density of the relative score by month for a given forecast horizon stratified by the one and four-week forecast horizon. c.) The average relative weighted interval score and absolute error for the surrogate model compared to the ensemble forecast by forecast horizon. d.) The density of the relative score by forecast location stratified by the one and four-week forecast horizon. The dashed line on all plots indicates when the ensemble forecast is equivalent to the surrogate forecast. The vertical black lines on the y-axis give individual relative scores."}
p <- (
  (
    (
      plot_rwis_horizon / plot_rwis_by_month
    )  +
    plot_layout(heights = c(4, 5), guides = "collect")
  ) |
  (
    plot_relative_summary(relative_summary) +
    plot_rwis_location +
    plot_layout(heights = c(1, 7), guides = "collect")
  )
) +
  plot_layout(guides = "collect", widths = c(1, 1)) +
  plot_annotation(tag_levels = "a") &
  theme(
    legend.position = "bottom", legend.margin = margin(),
    legend.justification = "centre"
  )
suppressWarnings(print(p))
```

>>>>

Whilst there is robust and consistent evidence that multi-team ensemble forecasts provide reliable and performant forecasts across domains [@Reich2022-jo] they also have a range of downsides.
The most significant is their considerable resource cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
Forecasts from these ensembles can also be hard to interpret as it is difficult to link them to the underlying model assumptions.
Additional issues with multi-team collaborations include a lack of feedback to those contributing forecasts that would allow them to improve their forecast approaches, maintaining incentives for forecasters to continue to contribute [@Bosse2022-hm], and difficulty improving the quality of the ensemble [@Sherratt2022-ue].
Each of these issues may impact the long-term quality of the resulting ensemble forecasts and have implications for end-users.

The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic due to a pressing need for reliable forecasts and a perception that many publicly available forecasts were low quality. Over 2020 and 2021, teams established COVID-19 Forecasting Hubs covering the US [@Cramer2022-sp], Germany and Poland [@Bracher2021-wk], and Europe [@Sherratt2022-ue] (the latter two including authors of this study). All of these collaborations ensembled contributions from multiple independent teams using a similar approach and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting.
However, less progress has been made mitigating the downsides of these projects and improving access to the high quality and robust forecasts they seek to generate to geographies without coverage or for other infectious diseases.

In this work, we draw insights from ensemble forecasts produced and endorsed by these forecast hubs, as well as our own forecasting work, to propose and evaluate a "surrogate" forecast model.
This seeks to mitigate some of the downsides of recent COVID-19 ensembles forecasts by emulating their observed behaviour, using a minimal set of easily communicated and epidemiologically justifiable assumptions, and limited computational resources with an easily generalised implementation.
The primary aim of this approach is to help highlight the behaviour, and potential mechanisms behind this behaviour, of ensemble forecast widely considered the gold standard for COVID-19 forecasting.
In addition to this, we seek to provide a simple, and robust, forecasting system that can be easily reused by others, perhaps where a forecast hub ensemble is not available, and be used as a platform for future research both into improving understanding of ensemble behaviour and increased forecast performance.

In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysic [@Vernon2014-ov], emulation approaches have been used to circumvent resource requirement issues for complex models by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components.
However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and, for ensemble models in particular, have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
This means that, whilst non-parametric emulation has utility for allowing rapid exploration of the parameter space for epidemiological models [@Iskauskas2022-fz; @harles2022-ds], it may be less suitable for use in replicating epidemiological ensemble forecasts than the surrogate assumption driven framework we propose.

In order to understand the behaviour of the forecast hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their own forecast models and then use them to produce weekly forecast that they then submit to the central hub team. 
Each hub targets a range of metrics including test positive reported cases, repoorted deaths, and hospitalisations, has a specific geographic focus, and asks for weekly forecasts (using epidemiological weeks i.e. Sunday to Saturday) over a time horizon of a few weeks.
Typically, the ensemble forecast produced each week comprises at least 3 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@Sherratt2022-ue].
Observed data are available and updated daily, and most submitted forecasts use this dataset, along with potentially other sources of real-time information, to produce forecasts.
Here, we focus on reported cases and primarily on the European forecast hub but our observations hold, in our view, across COVID-19 forecast hubs and to a lesser degree targets.
In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF; @Bosse2022-hm], and human insight based forecast models that may also include elements of other methods [@Bosse2022-hm].
Real-time evaluation has shown that each of these classes of models may perform well depending on the context and specific implementation of the forecast model [@Bosse2022-hm].

We have contributed a range of forecasts to these forecast collaborations, generally focussed on semi-mechanistic statistical methods, and human insight-based forecasts.
Our forecast submissions have not clearly over- or under- performed other forecasts submitted to the forecasting hubs (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm; @Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue]).
The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
Based on our observations our forecasts have generally captured the current trend relatively well, but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns. 
Our previous methodology also requires significant computational resources, running for an hour on a Azure D v5-series 16-core machine, when producing forecasts for the European forecasting hub [@AzurePricing]. 
In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
We found that our human insight-based forecasts significantly outperformed our model-based forecasts on average and hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@Bosse2022-hm].

@@

Based on these insights, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] with the aim of producing ensemble-like performance without sacrificing interpretability.
The first simplification we make is to model only weekly data, rather than using daily data and then aggregating. This mitigates the impact of daily reporting artefacts. It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain significant weight in the model.
This leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.

@@

The only observation for which we do not make an adaptation is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion.
This choice is motivated by our belief that the underlying transmission process is an exponential discrete one and therefore a count error model with a log link function, where variance is linked to the mean, is a sensible choice. 
We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations [@bracherEvaluatingEpidemicForecasts2021].

@@

```{r assemble-eval, fig.height = 12, fig.width = 12, out.width = "95%", fig.cap = "Relative weighted interval score by location, horizon, and forecast date for the surrogate forecast model compared to the ensemble forecast model on the log scale. a.) The density of the relative score by horizon. Horizontal black lines give the 5\\%, 35\\%, 65\\%, and 95\\% quantiles. b.) The density of the relative score by month for a given forecast horizon stratified by the one and four-week forecast horizon. c.) The average relative weighted interval score and absolute error for the surrogate model compared to the ensemble forecast by forecast horizon. d.) The density of the relative score by forecast location stratified by the one and four-week forecast horizon. The dashed line on all plots indicates when the ensemble forecast is equivalent to the surrogate forecast. The vertical black lines on the y-axis give individual relative scores."}
p <- (
  (
    (
      plot_rwis_horizon / plot_rwis_by_month
    )  +
    plot_layout(heights = c(4, 5), guides = "collect")
  ) |
  (
    (
      plot_relative_summary(relative_summary) +
      plot_layout(guides = "keep")
    )
    plot_rwis_location +
    plot_layout(heights = c(1, 7), guides = "collect")
  )
) +
  plot_layout(guides = "collect", widths = c(1, 1)) +
  plot_annotation(tag_levels = "a") &
  theme(
    legend.position = "bottom", legend.margin = margin(),
    legend.justification = "centre"
  )
suppressWarnings(print(p))
```
